Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-11 paper/paper.tex
update

\subsection{Sampling extremely large configuration space}
In many times, the number of configurations is too large to fit into any type of storage.
Then instead of using the set algebra, we can use a binary sum-product tree as a compact representation of the results.
We use a quadruple $(type, data, left, right)$ to represent a node in the tree, where $type$ is one of
$\{\texttt{LEAF}, \texttt{ZERO}, \texttt{SUM}, \texttt{PROD}\}$, $data$ is a bit string as the content is a \texttt{LEAF} node,
$left$ and $right$ are left and right operands of \texttt{SUM} and \texttt{PROD} nodes.

\begin{equation}
\eqname{SUMPROD}
\begin{split}
    s \oplus t &= (\texttt{SUM}, /, s, t)\\
    s \odot t &= (\texttt{PROD}, /, s, t)\\
    \mymathbb{0} &= (\texttt{ZERO}, /, /, /)\\
    \mymathbb{1} &= (\texttt{LEAF}, 0^{\otimes |V|}, /, /).
\end{split}\label{eq:expr}
\end{equation}
The vertex tensor and edge tensor become
\begin{equation}
    W^{\rm SUMPROD}((\texttt{LEAF},\boldsymbol{e}_i,/,/)) = \left(\begin{matrix}
        \mymathbb{1} \\
        (\texttt{LEAF},\boldsymbol{e}_i,/,/)
    \end{matrix}\right),   
    \qquad
        B^{\rm SUMPROD} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

This algebra is a commutative semiring because we define the equivalence of two sum-product trees by comparing their expanded (using \Eq{eq:set}) forms, rather than their storages.
Although one is probably unable to enumerate the configurations represented by this tree due to the extremely large configuration space size,
but he can directly sample configurations from this tree.
This is because once we have this sum-product representation of the configuration space,
one can compute the number of configurations in each branch rigorously and efficiently.
One can generate a random number and descend this expression tree to left or right sibling with probability decided by the size of each branch while doing the computation specified by the first field.

>>>>

\subsection{Sampling extremely large configuration space}
When the graph size goes large, the configurations might be impossible to fit into any type of storage.
Then instead of using the set algebra, one can use a binary sum-product tree as a compact representation of the configurations.
We use a quadruple $(type, data, left, right)$ to represent a node in the tree, where $type$ is one of
$\{\texttt{LEAF}, \texttt{ZERO}, \texttt{SUM}, \texttt{PROD}\}$, $data$ is a bit string as the content in a \texttt{LEAF} node,
$left$ and $right$ are left and right operands of \texttt{SUM} and \texttt{PROD} nodes.

\begin{equation}
\eqname{SPT}
\begin{split}
    s \oplus t &= (\texttt{SUM}, /, s, t)\\
    s \odot t &= (\texttt{PROD}, /, s, t)\\
    \mymathbb{0} &= (\texttt{ZERO}, /, /, /)\\
    \mymathbb{1} &= (\texttt{LEAF}, 0^{\otimes |V|}, /, /).
\end{split}\label{eq:expr}
\end{equation}
The vertex tensor and edge tensor become
\begin{equation}
    W^{\rm SPT}((\texttt{LEAF},\boldsymbol{e}_i,/,/)) = \left(\begin{matrix}
        \mymathbb{1} \\
        (\texttt{LEAF},\boldsymbol{e}_i,/,/)
    \end{matrix}\right),   
    \qquad
        B^{\rm SPT} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

This algebra is a commutative semiring because we define the equivalence of two sum-product trees by comparing their expanded (using \Eq{eq:set}) forms, rather than their storages.
Although one is probably unable to enumerate the configurations represented by this tree due to the extremely large configuration space size,
but he can directly sample configurations from this tree.
This is because once we have this sum-product representation of the configuration space,
one can compute the number of configurations in each branch rigorously and efficiently.
To generate samples, one just descends this sum product tree to left or right sibling with probability decided by the size of each branch while doing the computation specified by the first field.
